\section{Implementation Patterns and Caveats}

\subsection{Proper Usage of \texttt{lastDAprocessed} and \texttt{emitDA}}
\label{sec:lastDAprocessed}

\texttt{lastDAprocessed()} is a built-in method that helps you clean
up after a dialogue act has been dealt with. You usually want to call
it in your \texttt{propose} block, because when the block is executed
that means that the dialogue act has been processed. The method's
effect is to set an internal timestamp at the moment it has been
called, which affects the return value of \texttt{lastDA()}:
\texttt{lastDA()} will only return a dialogue act if it has been sent
after the the point in time specified by the \texttt{lastDAprocessed}
timestamp.

Be aware this also means that if the statement you execute in your
\texttt{propose} block is\\ \texttt{lastDAprocessed();}, all following
calls to \texttt{lastDA()} will evaluate to an empty dialogue
act. Thus, using expressions like \texttt{theme=lastDA().theme} in an
\texttt{emitDA} are strongly discouraged, because they will fail if
the \texttt{emitDA} is used after calling the cleanup method. There
is, however, good reason to not move the \texttt{lastDAprocessed()} to
the very end of your proposal, as proposals are executed in a separate
thread and your \vonda rules are (re-)evaluated in parallel. This
might, in rare cases where your proposal code takes more time to
process (for one possible reason, see \ref{sec:emitDA}), lead to your
system generating and executing new proposals based on the ''old''
dialogue act, thus responding more than once to one input.

\subsection{A Few Words About \texttt{emitDA} and \texttt{createBehaviour}} \label{sec:emitDA}

There is a feature to the \texttt{emitDA} method which has not been mentioned in section \ref{sec:caret}, but might become important in your specific application.

\texttt{emitDA} actually only is a wrapper method which uses the given
dialogue act to create a behaviour, which is the actual thing being
sent to the communication hub. \texttt{createBehaviour} wants to be
passed a delay parameter, which specifies the amount of time the
???communication thread???  \todo{AW: Is it the communication thread,
  or how can we call it?}  should be paused after emitting the given
behaviour. This might be important to your application if you use for
example TTS and want to delay the next utterance until playing the
current one has been finished.  Normal \texttt{emitDA} sets the delay
to \texttt{Behaviour.DEFAULT\_DELAY}, which by default is zero, but
you can also call \texttt{emitDA(delay, dialogueAct)} to directly
specify a delay, or even override \texttt{createBehaviour} to perform
a more complex computation of the delay time, e.g. to adopt to text
lenght * speed of your TTS voice.

\textbf{Attention!} Once you are doing this, make sure that you use
\texttt{lastDAprocessed()} early in your \texttt{propose} block as
suggested in \ref{sec:lastDAprocessed}. If you don't and the thread
the proposal is executed in is delayed long enough, new proposals will
be generated based on the old dialogue act and your agent might end up
saying things twice.

\subsection{Waiting for a User's Answer in a Conversation}

It's not very polite to be talking all the time without letting the
interlocutor say something themselves. Particularly, you'll want to
make sure that once the system asked a question, it will at least wait
for some time before going on, to give the user a chance to answer. To
this end, you can use the pre-built \texttt{waitingForResponse}
method, which returns true if the system was the last one to speak and
the dialogue act it uttered was a question or a request.

\subsection{Volatile variables in rule files and how to keep information between evaluation cycles} \label{sec:volatile}

During run-time, only one object of the top-level generated class
permanently exists. For all \texttt{include}d rule files, only
temporary objects are created, which live long enough to do all
relevant rule evaluations. That means that they have access to all
values of the top-level rule file and other rule files above them that
have been created up to the point of the rule file inclusion.

This means that whenever the \vonda rules are executed in a new cycle,
they are executed in a ``clean state'' where all variables you
previously set in the rule file have been reset. The only exception
from this are variables which are either located in your custom (Java)
Agent class or in your top-level rule file.

Thus: Always keep in mind that only variables defined in the Java
Agent instance or in the top-level rule file are persistent,
everything else is volatile! If you want to keep values between
executions, store them in a top-level variable or in the database.

\section{Advanced Features}

\subsection{Connecting to a second HFC Server} \label{sec:2ndHfc}
% e.g., like for PAL, if loading (a passive part of) the database takes too much time to perform it for every test
In the standard setup, your \vonda project uses one HFC server that on starting your system loads all the information from your ontology and that receives your new database entries and modifications.

However, there might be cases where this approach is not what you want. If your project uses a big database with static information, that you use but do not need to write to, you might prefer to not start the server anew each time you start your system, as this might consume time.


In this case, there is another solution: you can start your HFC server remotely and then connect to it in your \vonda agent.

On a linux machine, you can run a server by executing the following lines:
\todo{TODO: add hfc server start script}

To connect your \vonda agent to a local server you just need to add the following code, where \texttt{port} is the port you started it on and \texttt{myProxy} is the RdfProxy instance you can post your queries to.

\begin{center}
	\begin{lstlisting}[language=Java]
	myClient = RPCFactory.createSyncClient(HfcDbService.Client.class,
	"localhost", port);
	_myProxy = new RdfProxy(new ClientAdapter(myClient));
	\end{lstlisting}
\end{center}

This additional server does of course not have the same status as the innate HFC proxy, as you only connect to it at run-, not at compile-time. You can query it for information as described in \ref{sec:hfc_usage}, but classes from this database will not be recognized in the rudi code and you cannot write to it \todo{or can you? Try?!}.

\section{Troubleshooting: Typical Problems}

\begin{itemize}

	\item \textbf{The execution of my \texttt{propose} or \texttt{timeout} block does not have the effect I expected}\\
	Are you using any variables inside that block whose contents are changed by other parts of your code after the block has been issued?

	\item \textbf{The fields of the dialogue act sent by \texttt{emitDA} in my Proposal do not contain the values they should according to my conditions}\\
	Check whether you have ''buffered'' those values in final variables before issuing the Proposal and are referring to those. Using \texttt{lastDA} directly in the \texttt{propose} block is dangerous because it might already contain the next dialogue act (or none at all).

	\item \textbf{I get a NullPointerException in my Proposal}\\
	Please check whether all the variables you're using in the \texttt{propose} block are final and can't be changed by someone else between the time the Proposal is registered and the time it is executed. Also make sure you're not trying to read the fields of \texttt{lastDA()} after you called \texttt{lastDAprocessed}.

	\item \textbf{My system seems to execute the same Proposal multiple times}\\
	Make sure that in your \texttt{propose} block, you are calling \texttt{lastDAprocessed} as described in \ref{sec:lastDAprocessed} and also ''resetting'' everything else that triggers the rule to be executed.

	\item \textbf{The variable that I use in my rules for storing information does not have the contents it is supposed to}\\
	Be aware that variables not defined in the top-level rule file are not persistent between rule evaluation cycles, thus you should not store information there which you want to keep (see section \ref{sec:volatile}).

	\item \textbf{When compiling, I get the warning \texttt{''base name x can be one of y, z, please resolve manually''}}\\
	It seems your ontology is redefining an existing class. Refer to section \ref{sec:nsAmbigue} to see how to resolve the problem.

\end{itemize}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "userguide"
%%% End:
